---
title: "Computing zones of influence on GRASS GIS through the oneimpact package in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Computing zones of influence on GRASS GIS through the oneimpact package in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The computation of zones of influence (ZoI) in useful to assess whether the effect of 
a given anthropogenic infrastructure, disturbance, or landscape variable affect
an ecological response variable beyond its spatial delimitation. Computing zones of 
influence for large maps with fine resolution though, is tricky because of computing and
memory limitations in processing raster maps in R. To deal with that, the `oneimpact`
package offers tools to compute the ZoI in [GRASS GIS](https://grass.osgeo.org/), a
powerful, open-source , using the package [`rgrass7`](https://rsbivand.github.io/rgrass/) 
to link R and GRASS. Here we provide an example and some comments on how to
compute ZoI metrics in GRASS GIS via R.

For an introduction to the zone of influence approach in R, look at the 
[Getting started](https://ninanor.github.io/oneimpact/articles/oneimpact.html) and
[Computing zones of influence](https://ninanor.github.io/oneimpact/articles/computing_zoi.html)
introdutory vignettes.

# Overview

To compute ZoI metrics in GRASS GIS,
we can use the same function used to compute ZoI in R, `calc_zoi()` (or 
`calc_zoi_nearest()` and `calc_zoi_cumulative` for only the nearest and cumulative
ZoI metrics), adding a parameter `where = "GRASS"`. However, the computation in GRASS GIS 
is slighly different from doing the same in R. Here are some key differences:

- raster processing is made in GRASS GIS, so it requires an active connection between
the R session and a GRASS GIS location and mapset;
- the input raster maps representing infrastructure or disturbance might be loaded into
the GRASS GIS location/mapset R is connected to. This might be done through the
`rgrass7` package;
- once the input raster maps are loaded in GRASS, the input to the `calc_zoi()`
function is not a raster object in R but a string with the name of the map in the
GRASS mapset;
- after processing, ZoI maps are saved within the GRASS GIS location and mapset.
To retrieve them, one must use the `rgrass7` pacakge to import the maps into R
or outside to a working folder.

Below we show how to do these steps to compute the ZoI and provide a few comments.

# Establishing a connection between R and GRASS

## Downloading and installing GRASS GIS

The first requirement to compute ZoI metrics in GRASS is to have GRASS GIS installed.
GRASS can be freely downloaded and installed in most operational systems 
(Windows, Linux, MacOS). For more information on installations, please see the official
[GRASS Downloads webpage](https://grass.osgeo.org/download/). The website also
provides [helpful manuals, tutorials, and videos](https://grass.osgeo.org/learn/) 
and various languages for first time and experienced users.

## Connecting R to GRASS GIS

GRASS GIS has a slighly different project organization, if compared to other
popular GIS software such as ArcGIS and QGIS. GRASS projects are called **Locations**,
which are defined by a single [geodetic datum](https://en.wikipedia.org/wiki/Geodetic_datum) 
and [coordinate reference system (CRS)](https://rspatial.org/terra/spatial/6-crs.html).
This means that all maps of all types (e.g. vector, raster) within a location
have the very same CRS, what assures all data is processed correctly. Each location
has a main folder called **PERMANENT** and might have multiple other folders; these folders
are called **Mapsets**. See 
[this tutorial](https://neteler.gitlab.io/grass-gis-analysis/01_grass-gis-basics/) 
for more information on the GRASS database and location structure.

### Running GRASS from R

When connecting R to GRASS, R might be connected to a given GRASS location and mapset
through the function `rgrass7::initGRASS()`. One must provide to this call the folder where the 
GRASS GIS executable files are found (parameter `gisBase`), as well as the folder where
the project is located (parameter `gisDbase`), the location name (parameter `location`)
and the mapset name (parameter `mapset`). The function might be used to connect to an 
already existing location or create a new one, if it does not exist.

For illustration, here we create a new location to work on the `oneimpact` examples.
To do so, we use as a reference (for defining the datum and CRS of the GRASS location)
the map of number of private tourist cabins in a sample area in Southern Norway, 
stored in the `oneimpact` package (see `?sample_area_cabins_count.tif`
for more information). We first load this data set.

```{r load_cabins}
# load packages
library(oneimpact)
library(rgrass7) # for connecting R and GRASS
library(terra) # for geodata in R

# load cabins raster data into R
(f <- system.file("raster/sample_area_cabins_count.tif", package = "oneimpact"))
cabins_count <- terra::rast(f)
```

Next, we need to know what is the path of the folder where GRASS GIS executable
files are located. The following command works in Linux, MacOS, and within the OSGeo4W
in Windows:

```{r where_grass}
# For linux or within OSGeo4W shell
grassdir <- system("grass78 --config path", intern = TRUE)
grassdir
```

Now we can create a GRASS GIS location and connect our R session to it:

```{r create_grassdb, message=FALSE, warning=FALSE, error=FALSE}
gisDB <- "." # create location and mapset in the working directory
loc <- "test_location/" # name of the location
ms <- "PERMANENT" # name of the mapset
# create grass db and connect
rgrass7::initGRASS(gisBase = grassdir,
                   SG = cabins_count, # use map to define location projection
                   home = tempdir(),
                   gisDbase = gisDB,
                   location = loc,
                   mapset = ms,
                   override = TRUE, )
```

Notice that, in the code above, a few additional parameters were filled. First, 
the `SG = cabins_count` parameter defined the map of cabins as the reference to 
define the CRS of the created location. Second, we created a location in a temporary
folder using `home = tempdir()`. While this is ok for tests and demonstration purposes like 
this tutorial, we do not recommend you to use it when processing maps for other purposes,
since the data might be lost after the connection to GRASS is closed. 

As a proof to check if the connection between R and GRASS is active, the `initGRASS()`
function shows a series of information about the grass database and location. The
same might be checked at any time by using the `rgrass7::gmeta()` function:

```{r check_grass, eval=FALSE}
# check grass connection
rgrass7::gmeta()
```

This command shows all the information on the grass database, location, mapset, and
reference system. Besides, it provides information about the GRASS working computational region
(number of `rows` and `columns` of the region; `north`, `south`, `west`, and `east` boundaries of 
the region extent; and region resolution in the north-south and east-west directions,
`nsres` and `ewres`). Differently from other GIS, GRASS works based on computational regions
to perform map processing. This means that, regardless of the extent and resolution of the input
maps, it is possible to set a specific desired extent and resolution as the output for most
of the GRASS GIS tools. More information on GRASS computational regions can be found 
[here](https://neteler.gitlab.io/grass-gis-analysis/01_grass-gis-basics/#concept-of-computational-region).

### Running R within GRASS

Another option is to open a GRASS GIS session and run R within the GRASS shell.
Too know more about this option, please look at these GRASS GIS 
[wiki](https://grasswiki.osgeo.org/wiki/R_statistics/rgrass7).

## Notes in GRASS installation and connections on Windows

It is possible to install GRASS on Windows operational systems in two ways, and it might
be a little tricky to connect R and GRASS. 

### Standalone installation

The first option is to download a [standalone installer](https://grass.osgeo.org/download/windows/) 
and install a given GRASS GIS
version. If you do so, the code above should work, except the command 
`system("grass78 --config path", intern = TRUE)`. In this case, the `grassdir` object
with the path to the executable GRASS files should look like:

```{r grass_windows, eval=FALSE}
# If you used the standalone installer in Windows
grassdir <- "C:\Programs\GRASS GIS 7.8" # Correct if the path is different
```

### Installation through OSGeo4W

Another option on Windows is to use [OSGeo4W](https://grass.osgeo.org/download/windows/). 
OSGeo4W is an installer for a broad set of open source geospatial software packages including GRASS GIS 
as well as many other packages (QGIS, GDAL/OGR, and more). Here it is possible to choose which version of 
GRASS GIS to install and choose many other installation setting besides the default.

When GRASS is installed through OSGeo4W, though, R must be opened from the OSGeo4W shell
for the connection between R and GRASS to work through the `rgrass7` package. See the
documentation for the `rgrass7::initGRASS()` function for more information.

# Computing the zones of influence

## Loading input maps into GRASS

Before computing the ZoI metrics, the input infrastructure or disturbance maps must
be loaded into GRASS. If the maps are already open in our R session, it is possible
to write them into our GRASS mapset using the `rgrass7::write_RAST()` function:

```{r write_raster1}
# write raster into GRASS
rgrass7::write_RAST(cabins_count, 
                    vname = "cabins_count", # name of the map in GRASS
                    flags = c("o", "quiet"))
```

In some cases, there might be small, negligible differences between the GRASS location
CRS and the the input map CRS. In such cases where this difference might be ignored,
the parameter `flags = "o"` might be added to ignore the differences (look 
[here](https://grass.osgeo.org/grass82/manuals/r.in.gdal.html) for more details on
the `-o` flag). Likewise, if the map already exists in the GRASS location and we want
to overwrite it, we must add the `overwrite = TRUE` parameter to the function.

Another way of loading a map into a GRASS location and mapset directly from a folder 
in your computer, without the need to pre-load it into R, is to use the 
[`r.in.gdal`](https://grass.osgeo.org/grass82/manuals/r.out.gdal.html) or 
[`r.import`](https://grass.osgeo.org/grass82/manuals/r.import.html) GRASS tools through 
the `rgrass7::execGRASS()` function. The function `execGRASS()` provides a way of 
running GRASS modules and addons directly, as if we were running them through the GRASS
console.

In this case, the following command would also work:

```{r write_raster2, eval=FALSE}
# get path to the raster file
f <- system.file("raster/sample_area_cabins_count.tif", package = "oneimpact")
f

# load this file to the GRASS mapset
rgrass7::execGRASS("r.in.gdal", input = f, 
                   output = "cabins_count", # name of the map in GRASS
                   flags = c("overwrite", "o"))
```


## Computing the ZoI

Now, computing the ZoI is very similar to the procedure in R with the `calc_zoi()`
function. We will here
compute both the ZoI of the nearest feature and the cumutive ZoI for cabins
using a ZoI with a linear decay (i.e. Bartlett) shape and radii of 500 and 1000 m. 
Importantly, here the main input `x` for the `calc_zoi()` funcion is a string
with the name of the map within the GRASS mapset (or a vector strings for multiple maps).
To inform R that we wish to make processing in GRASS, we must set `where = "GRASS"`:

```{r compute_zoi}
# compute zoi
zoi_metrics <- calc_zoi(x = "cabins_count", # string as input
                        radius = c(500, 1000),
                        type = "bartlett",
                        zeroAsNA = TRUE,
                        where = "GRASS", 
                        g_overwrite = TRUE)

# check the output
zoi_metrics
```

Notice that, differently from when `calc_zoi()` is run entirely in R, the output object `zoi_metrics`
is a vector of strings specifying the name of the output maps computed in GRASS. Below we
show how to retrieve these maps into R or export them from GRASS into a local folder.

Notice also that the `g_overwrite` parameter was used to guarantee that, if there are already
maps with similar names in the GRASS GIS mapset, they will be overwritten. Other parameters
to control how `calc_zoi()` is computed in GRASS are available, all following starting 
with `g_`. For instance, the name of the output map (or pattern in the output names, when there
is more than one map or ZoI radius) might be determined using the `g_output_map_name` parameter.
If no parameter is specified, the input map name is used, together with the ZoI metric, shape, and radius,
to create the name of the output maps. For more information on GRASS GIS options for 
ZoI computation, see the documentation for `calc_zoi_nearest()` and `calc_zoi_cumulative()`.

# Retrieving outputs from GRASS

Once the ZoI maps are calculated, we can retrieve them to R using the function
`rgrass7::read_RAST()`. By default, this function reads the rasters as a 
`SpatialGridDataFrame` object from the `sp` package. Here we prefer to specify
`return_format = "terra"`, in case which the output is a *terra* `SpatRaster` object.

```{r retrieve_maps}
# retireve maps
cabins_zoi_cumulative <- rgrass7::read_RAST(zoi_metrics, 
                                            return_format = "terra",
                                            flags = "quiet")
cabins_zoi_cumulative
```

```{r fix_raster, eval=TRUE, include=FALSE}
# For some reason the nearest maps are being read/approximated to integers
# we fix it here, later on we look for the issue
# probably a terra package issue, or a rgrass issue
cabins_zoi_cumulative[[1]][] <- as.numeric(cabins_zoi_cumulative[[1]][])
cabins_zoi_cumulative[[2]][] <- as.numeric(cabins_zoi_cumulative[[2]][])
cabins_zoi_cumulative[[3]][] <- as.numeric(cabins_zoi_cumulative[[3]][])
cabins_zoi_cumulative[[4]][] <- as.numeric(cabins_zoi_cumulative[[4]][])
```


Now we can plot the ZoI metrics computed in GRASS:

```{r}
# plot ZoI metrics
terra::plot(cabins_zoi_cumulative)
```

From this point, the `oneimpact` vignette [Getting started](https://ninanor.github.io/oneimpact/articles/oneimpact.html) shows how it is
possible to annotate biological data with the ZoI variables computed for different
variables to prepare data for a multi-scale analysis on the effects of infrastructure
and anthropogenic disturbance on biological responses (see also Niebuhr et al, 
*submitted*).

# Additional resources

## Different algorithms to compute the cumulative zone of influence

In GRASS GIS, different algorithms might be used for the computation of 
neighborhood analyses, which are the basis for the cumulative ZoI calculation.
The following GRASS tools are available to be used in `calc_zoi()` in GRASS:
`r.resamp.filter`, `r.mfilter`, or `r.neighbors`. The module to be used is
controlled by the parameter `g_module`. Each of these algorithms provide different
capabilities and flexibility:

- `r.resamp.filter` seems to be the fastest one
in most cases, but has less flexibility in the choice of the zone of influence
shapes. The algorithm calculates the weighted density of features, which
might be rescaled to the cumulative ZoI if the appropriate scaling factor
(calculated from the weight matrix) is provided. Currently only the
filters `type = "bartlett"` and `type = "box"` are implemented in
`oneimpact`. More
information about the algorithm
[here](https://grass.osgeo.org/grass82/manuals/r.resamp.filter.html).
- `r.mfilter` (the default tools in the `calc_zoi()` functions) 
is slower than `r.resamp.filter` but much faster than
`r.neighbors`, and allows a flexible choice of the shape of the zone of
influence (the wight matrix shape). `r.mfilter` is then the most indicated
in terms of a balance between flexibility in the choice of the ZoI shape
and computation efficiency.
The only inconvenient of `r.mfilter` is that it
creates an edge effect with no information in the outer cells of a raster
(the number of cells correspond to `radius` or half the size of the weight
matrix), so if it is used the users should add a buffer area
\eqn{ge radius} around the input raster map, to avoid such edge effects.
See \url{https://github.com/OSGeo/grass/issues/2184} for more details about 
this issue, and [this link](https://grass.osgeo.org/grass82/manuals/r.mfilter.html)
for more information on this tool.
- `r.neighbors` is considerably slower than the other algorithms (from 10 to
100 times), but allows a flexible choice of the ZoI shape. Contrary to
`r.resamp.filter` and `r.mfilter`, which can only perform a sum of pixel
values weighted by the input filter or ZoI, `r.neighbors` might
calculate many other statistical summaries within the window of analysis,
such as mean, median, standard deviation etc. In principle, though, the sum
is enough to compute the cumulative ZoI as formulated in Niebuhr et al. (*submitted*)
and proposed here.

## Different measures of distance to compute the zone of influence of the nearest feature

The metrics of ZoI of the nearest feature are based on maps of distance form each 
pixel in the landscape to the nearest feature. They are computed in GRASS through the
[r.grow.distance](https://grass.osgeo.org/grass78/manuals/r.grow.distance.html)
tool. This tool allows the computation of the distance to the nearest feature based
on different distance metrics:  "euclidean", "geodesic", "squared", "maximum", "manhattan".
This allows extra flexibility and might be controlled in the `calc_zoi()` functions
through the `g_dist_metric` parameter. If `g_dist_metric = "geodesic"` and one uses 
the option `flags = "m"`, for instance, the computation of the distance is made in 
meters even if the map is in geographical coordinate system (Latitude-Longitude).
Find more information at the GRASS GIS documentation for this function 
[here](https://grass.osgeo.org/grass82/manuals/r.grow.distance.html).

## Additional ancillary functions

Additional examples for the computation of the ZoI in GRASS can be 
found in the documentation of the `calc_zoi_nearest()`
and `calc_zoi_cumulative()` functions. The `oneimpact` package also provides a 
few other ancillary functions to help processing maps within GRASS:

- The function `grass_binarize()` transforms continuous maps into binary maps
based on `breaks` to define discrete categories. The output consists of maps with
1 for values above the `break` and 0 for values below it. It might be used for 
preparing inputs to compute the ZoI based on continuous layers.
- The function `grass_v2rast_count()` rasterizes vectors within GRASS 
by counting the number of features in each pixel. This is usefull for large maps
which are hard to be rasterized within R. See the 
[Computing zones of influence](https://ninanor.github.io/oneimpact/articles/computing_zoi.html)
vignette for a reasoning about where this might be useful.

# References

Niebuhr, B. B., van Moorter, B., Stien, A., Tveraa, T., Strand, O., Langeland, K.,
Alam, M., Skarin, A., & Panzacchi, M. Estimating the cumulative impact and zone of
influence of anthropogenic infrastructure on biodiversity. *Submitted manuscript*.
